---
title: "Data Science for Public Policy"
subtitle: "Assignment 04"
author: "Priscila Stisman - prs62 & Maleeha Hameed - mh2203"
execute: 
  warning: False
format: 
  html: 
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

## **Exercise 3.** 

**Try to git add your data file. What message do you receive in the
terminal?**

#hint: Use -f if you really want to add them. #hint: Turn this message
off by running #hint: "git config advice.addIgnoredFile false" #The
`echo: false` option disables the printing of code (only output is
displayed).

```{r}
# Library setup
library (tidyverse)
library (dplyr)
library(ggplot2)
library(ggthemes)
library(readr)
library (haven)
```

## **Exercise 4** 

Our visualizations are based on the Quality of Government (QoG)
Institute Basic Dataset 2023 (add citation). This dataset contains \[add
more details later\].

```{r}
# Read data
data <- read_dta("Data/idm_qog.dta")

```

#### Graph 1: People in countries with higher GDP per capita live longer 

This scatter plot shows the correlation between the log of the GDP per
capita (in US dollars) and the life expectancy in years for XXX
countries. Except for a few outliers with a higher life expectancy for a
lower GDP per capita, we find a linear correlation (denoted by a line of
best fit) between countries with higher GDP per capita and their life
expectancy in years. This graph confirms the success story of many
developed countries, globally, where people live longer than 75 years of
age, on average, given higher annual earnings.

```{r}
# Graph 1

data %>%
  ggplot() +
    geom_point(aes(x = log(gle_cgdpc), y =wdi_lifexp)) + 
  geom_smooth(aes(x = log(gle_cgdpc), y = wdi_lifexp), method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
   xlab("Log GDP per Capita") +  # Set x-axis label
  ylab("Life Expectancy")      # Set y-axis label
```

#### Graph 2: Countries with higher GDP per capita contribute more to climate change 

A higher GDP per capita does not always tell an optimistic story! With
climate change increasingly becoming a pressing global policy issue, we
want to draw attention to the correlation between countries with a
higher GDP per capita and their contribution to a higher carbon
footprint of production per year. We use the color scale gradient from
light blue to indigo and the size of the circle to show the variation in
the carbon footprint of production.

```{r}
# Graph 2

ggplot(data, aes(x = log(gle_cgdpc), y = 0.1, size = ef_carbp, color = ef_carbp), na.rm = TRUE) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10)) +
  scale_color_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Scatter Plot: Carbon footprint of production and GDP",
       x = "Log GDP",
       y = "",
       size = "Carbon footprint of production",
       color = "Carbon footprint of production") +
  guides(size = guide_legend(title = "Carbon footprint of production"))


```

#### Graph 3: Cross region variation in how long people live 

Except for Sub Saharan Africa, the rest of the 9 regions of the world
have an average life expectancy of greater than 60 years. North America
and Western Europe, where most developed countries are located, fare the
highest on the average life expectancy. We do not observe much variation
in life expectancy over time across the different regions as most of the
variation can be observed at the inter-country level. This graph
reinforces the findings from Graph 1.

```{r}
# Graph 3
library(ggplot2)

# Assuming your data frame is named data
ggplot(data, aes(x = factor(ht_region), y = wdi_lifexp, fill = wdi_lifexp)) +
  stat_summary(fun = "mean", geom = "bar", fill = "lightblue", alpha = 0.7) +
  labs(title = "Average Life Expectancy by Region",
       x = "Region",
       y = "Average Life Expectancy") +
  theme_minimal() +
  scale_x_discrete(labels = c("Eastern Europe and post Soviet Union", "Latin America", "North Africa & the Middle East", "Sub-Saharan Africa", "Western Europe and North America", "East Asia", "South-East Asia", "South Asia", "The Pacific", "The Caribbean"))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

#### Graph 4: Women make less than 50% of the labor force globally 

We take advantage of the time-series data to capture another important
policy issue – distribution of female labor force participation across
all the countries between 2000 and 2021. The line graph shows a steep
increase in the average female labor force participation with a sharp
drop in 2020, at the onset of COVID-19. We observe an upward trend after
2020 until 2021, however, the average female labor force participation
does not reach the pre-COVID high. Unfortunately, despite multiple
reforms to improve female labor force participation globally, the share
of average female labor force participation stands at less than 41.5% of
the total average labor force participation.

```{r}
# Graph 4

average_female_labor_participation <- data %>%
  group_by(year) %>%
  summarize(avg_female_labor_participation = mean(wdi_lfpf, na.rm = TRUE))

# Plotting the line plot
ggplot(average_female_labor_participation, aes(x = year, y = avg_female_labor_participation)) +
  geom_line() +
  labs(title = "Average Female Labor Force Participation by Year",
       x = "Year",
       y = "Average Female Labor Force Participation")

```
